Viscous Transport in Eroding Porous Media

### 2019-10-15 docs/transport/erosion_transport.tex
working on formatting and simplifying

We consider trajectories in an open channel so that they can be compared
to trajectories in eroded geometries with large permeability.  The
trajectory originating at $(s_1(0),s_2(0))$ is
\begin{align}
  s_1(t) = (1-s_2(0)^2)t + s_1(0), \quad
  s_2(t) = s_2(0).
\end{align}
Averaging over $s_2(0) \in [-1,1]$, the first and second ensemble
moments are
\begin{align}
  \langle \lambda \rangle (t) = \frac{2}{3}t, \quad 
    \sigma_\lambda^2(t) = \frac{4}{45}t^2.
\end{align}
In section~\ref{sec:results}, we compare the second moment in an open
channel with the trajectories in an eroded geometry.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The tortuosity is a dimensionless number that quantifies the amount of
twisting of streamlines. Unlike the dispersion calculations, we do not
use reinsertion to form long trajectories.  The local tortuosity is
\begin{align}
  \tau(y_0) = \frac{\lambda(y_0)}{d}.
  \label{eqn:localTort}
\end{align}
Here the streamline originates on the inlet cross-section $x=x_0$ at
$(x_0,y_0)$, and its arclength, $\lambda(y_0)$, is calculated until the
streamline passes the parallel outlet cross-section $x = x_0 + d$.  In
this work, we consider streamlines originating at $x=-1$ and terminating
at $x=1$, so $d=2$.  However, other choices for the terminal point when
computing the tortuosity are sometime used~\citep{dud-koz-mat2011}.  The
hydraulic tortuosity is defined by taking the average over all points on
the inlet cross-section
\begin{align}
  T = \frac{1}{d}\frac{\displaystyle\int_{S}u_1(x_0,y_0)\lambda(y_0)\,dy_0}
  {\displaystyle\int_{S}u_1(x_0,y_0)\,dy_0},
  \label{eqn:tortuosity1}
\end{align} 
where $S$ is the inlet cross-section $x = -1$, $u_1(y_0)$ is the
$x$-component of the velocity at the initial point of the streamline,
and $d=2$ is the distance between the inlet and outlet.  Note that $T
\geq 1$, and $T=1$ only if no grains are present.

The tortuosity can also be computed with an area integral. Assuming that
the flow is incompressible and not re-entrant, meaning that all
streamlines connect the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent
to~\citep{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\citep{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in section~\ref{sec:results}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in porous media depends on the porosity, it also depends
on the placement of the grains.  In particular, grain placement affects
velocity scales~\citep{ali-par-wei-bre2017}, correlation
structures~\citep{leb-ded-dav-bou2007}, contaminant
transport~\citep{knu-car2005},
channelization~\citep{sie-ili-pri-riv-gua2019,berhanu2012shape}, and
pore network models~\citep{bry-kin-mel1993, bry-mel-cad1993,
bij-blu2006}. To characterize the grain placement, we compute
distributions of pore sizes between neighboring grains.  To define
neighboring grains, we form a Delaunay triangulation using nodes placed
at the center of each eroding grain and at a collection of points around
the boundary of the porous media.  Then, we say that two grains are
neighbors if their centers share an edge of the
triangulation~\citep{dea-qua-bir-jua2018}. The pores of an eroded
geometry are illustrated in figure~\ref{fig:Eroding100gap}(a). We do not
consider pores between eroding bodies and the solid wall $\Gamma$, so
some of the grains near the porous region boundary only have two
neighbors. Having defined the pore sizes, we plot its distribution in
figure~\ref{fig:Eroding100gap}(b) and compare it with the Weibull
distribution, a distribution used by others to characterize pore
sizes~\citep{ioa-cha1993}.  In section~\ref{sec:Eroding100}, we
investigate the effect of erosion on the pore size distribution.
\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The pore sizes between
neighboring grains. Two grains are said to be neighbors if they share an
edge of a Delanuay triangulation with nodes at the center of each
eroding grain. (b) The distribution of the pore sizes of an eroded
geometry. The black curve is the Weibull distribution with the same
first two moments as the data.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Consider a single eroding body close to a solid wall with the shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. \citet{mit-spa2017} performed a similar
three-dimensional experiment using a second-order quadrature method.
Their initial body is a sphere with its center located $1.5$ radii above
the solid wall.  We initialize the two-dimensional eroding body with
radius $r = 0.4$, and we conduct numerical experiments where the initial
distance between the grain and the solid wall is $h$, $h/2$, and $h/10$,
where $h = 2\upi r/N_\iin$.  If we used the trapezoid rule and required
an error that is comparable to the Barycentric quadrature rule, the body
with an initial distance of $h/10$ from the solid wall would require
$6,400$ discretization points, and the outer wall would require $50,000$
discretization points.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding grains with the Hagen-Poiseuille flow $\UU(\xx) =
U \left(1-y^2,0 \right)$.
%\begin{align}
%  \UU(\xx)=U \left[
%  \begin{array}{c}
%    1-y^2 \\ 0
%  \end{array}
%  \right],
%\end{align}
imposed on $\Gamma$. The flow rate $U$ is chosen so that the average
pressure drop from $x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$
once all the grains have vanished.  The vorticity and grain
configuration at four equispaced times are shown in
figure~\ref{fig:Eroding20vort}.  Initially, several of the grains are
closer to the outer wall than the $5h$ threshold required for the
trapezoid rule to achieve machine precision.  In particular, the
distance between bodies 1, 6, 13, and 15 and the outer wall is $1.3h$,
$2.9h$, $2.8h$, and $1.3h$, respectively, where $h$ is the arclength
spacing of the outer wall $\Gamma$.  In addition, the distance between
several pairs of eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and
14 \& 18, is too small to be resolve with the trapezoid rule.  By using
the Barycentric quadrature rule, the interaction between these
nearly-touching bodies is resolved to the desired accuracy, and erosion
can be simulated until all the bodies have vanished.

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies in a Hagen-Poiseuille flow.  The four snapshots are evenly spaced
in time, and the color is the fluid vorticity. In the fourth frame,
bodies 4, 12, and 17 have completely vanished, and body 19 has almost
completely eroded.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20densen}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as an exponential law  (dashed line) towards the
flow rate $U=1$ that occurs once all the bodies have eroded.}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from figure~\ref{fig:Eroding20vort}. The streamlines are
initially equispaced at $(-1,y)$, where $y \in (-1,1)$. The first five
snapshots are evenly spaced in time.  The bottom right frame is a zoom
in of the fifth snapshot, but with additional streamlines.  Since we use
the Barycentric quadrature rule and a fourth-order Runge-Kutta method,
streamlines that are very close to the grains are resolved.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of a porous geometry
that initially had 20 grains and has eroded to a porosity of 62.9\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1,y)$,
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ on the cross section $x = -1$.  The spatially
averaged tortuosity through the channel is $1.098$.  (c) The streamlines
resulting in the ten largest differences of local tortuosity between
neighboring streamlines.  Neighboring streamlines have the same color.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} The erosion of 20 nearly touching
bodies in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity. In addition to the
channels that develop between the bodies and the solid walls, erosion
leads to two main channels through the geometry---one in the top half
and one near the middle.}
\end{figure}

@@

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_50}
\caption{100 bodies, $\phi = 55.54\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_100}
\caption{99 bodies, $\phi = 62.98\%$}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_150}
\caption{94 bodies, $\phi = 72.22\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_200}
\caption{82 bodies, $\phi = 83.67\%$}
\end{subfigure}
\caption{\label{fig:Eroding100vort} The erosion of 100 nearly touching
grains in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity.  Because of the large
number of bodies, erosion creates many channels connecting the inlet to
the outlet.}
\end{figure}

In figure~\ref{fig:Eroding100Transport}(a), we plot the tortuosity as a
function of the porosity.  The initial geometry has a porosity of $\phi
= 50.09\%$ and the tortuosity is $T = 1.20$.  The tortuosity is computed
with both the length of the streamlines~\eqref{eqn:tortuosity1} (red
stars) and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding100tort}(c).
Again, the two tortuosity formulas give similar results.  For this
geometry, the tortuosity decreases monotonically at almost all
porosities.  However, the tortuosity undergoes a sudden increase near
the end of the simulation, and we have observed this behavior in other
examples.  The increase is caused by a single small body near the middle
of the channel being completely eroded.  While this results in
straighter streamlines, therefore reducing $\lambda$, the horizontal
flow, $u_1(y)$, increases since there is no longer a no-slip boundary,
and this increases the tortuosity.  We also compute the lines of best
fit using the porosity-tortuosity models~\eqref{eqn:tortuosityModels}.
The black dashed line in figure~\ref{fig:Eroding100Transport}(a) is the
line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2459},
\end{align}
with a root-mean-square error of $5.50 \times 10^{-3}$.  We note a
slightly better root-mean-square error of $5.20 \times 10^{-3}$ is
possible with the model
\begin{align}
  \widehat{T}(\phi) = 1 - 0.2631 \ln(\phi).
\end{align}
\todo[inline]{More inline candidates}

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity of a porous geometry
that initially had 100 grains and has eroded to a porosity of 62.98\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1, y)$,
normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.  Note
that this maximum velocity is about a order of magnitude smaller than
the 20 body example in figure~\ref{fig:Eroding20tort}.  (b) The local
tortuosity $\tau(y)$ on the cross section $x = -1$. The spatially
averaged tortuosity through the channel is $T = 1.12$. (c) The
trajectories of 200 tracers initialized at $x = -1$.  The
discontinuities in the local tortuosity correspond to nearby streamlines
whose lengths are significantly different because they pass opposite
sides of an eroding body. Compared to figure~\ref{fig:Eroding20tort},
this example has more small bodies, and this results in many more
discontinuities.}
\end{figure}

>>>>

We consider trajectories in an open channel so that they can be compared
to trajectories in eroded geometries with large permeability.  The
trajectories are $(s_1(t),s_2(t)) = ((1-s_2(0)^2)t + s_1(0), s_2(0))$.
%\begin{align}
%  s_1(t) = (1-s_2(0)^2)t + s_1(0), \quad
%  s_2(t) = s_2(0).
%\end{align}
Averaging over $s_2(0) \in [-1,1]$, the first and second ensemble
moments are
\begin{align}
  \langle \lambda \rangle (t) = \frac{2}{3}t, \quad 
    \sigma_\lambda^2(t) = \frac{4}{45}t^2.
\end{align}
In section~\ref{sec:results}, we compare the second moment in an open
channel with the trajectories in an eroded geometry.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The tortuosity is a dimensionless number that quantifies the amount of
twisting of streamlines. Unlike the dispersion calculations, we do not
use reinsertion to form long trajectories.  The local tortuosity is
\begin{align}
  \tau(y_0) = \frac{\lambda(y_0)}{d}.
  \label{eqn:localTort}
\end{align}
Here the streamline originates on the inlet cross-section $x=x_0$ at
$(x_0,y_0)$, and its arclength, $\lambda(y_0)$, is calculated until the
streamline passes the parallel outlet cross-section $x = x_0 + d$.  In
this work, we consider streamlines originating at $x=-1$ and terminating
at $x=1$, so $d=2$.  However, other choices for the terminal point when
computing the tortuosity are sometime used~\citep{dud-koz-mat2011}.  The
hydraulic tortuosity is defined by taking the average over all points on
the inlet cross-section
\begin{align}
  T = \frac{1}{d}\left(\int_{S}u_1(x_0,y_0)\lambda(y_0)\,dy_0 \right)
  \Bigg/
  \left(\int_{S}u_1(x_0,y_0)\,dy_0 \right),
%  T = \frac{1}{d}\frac{\displaystyle\int_{S}u_1(x_0,y_0)\lambda(y_0)\,dy_0}
%  {\displaystyle\int_{S}u_1(x_0,y_0)\,dy_0},
  \label{eqn:tortuosity1}
\end{align} 
where $S$ is the inlet cross-section $x = -1$, $u_1(y_0)$ is the
$x$-component of the velocity at the initial point of the streamline,
and $d=2$ is the distance between the inlet and outlet.  Note that $T
\geq 1$, and $T=1$ only if no grains are present.

The tortuosity can also be computed with an area integral. Assuming that
the flow is incompressible and not re-entrant, meaning that all
streamlines connect the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent
to~\citep{dud-koz-mat2011}
\begin{align}
  T = \left(\int_\Omega \|\uu(\yy)\|\, d\yy \right)
    \Bigg/
      \left(\int_\Omega u_1(\yy)\, d\yy \right),
%  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
%           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\citep{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in section~\ref{sec:results}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in porous media depends on the porosity, it also depends
on the placement of the grains.  In particular, grain placement affects
velocity scales~\citep{ali-par-wei-bre2017}, correlation
structures~\citep{leb-ded-dav-bou2007}, contaminant
transport~\citep{knu-car2005},
channelization~\citep{sie-ili-pri-riv-gua2019,berhanu2012shape}, and
pore network models~\citep{bry-kin-mel1993, bry-mel-cad1993,
bij-blu2006}. To characterize the grain placement, we compute
distributions of pore sizes between neighboring grains.  To define
neighboring grains, we form a Delaunay triangulation using nodes placed
at the center of each eroding grain and at a collection of points around
the boundary of the porous media.  Then, we say that two grains are
neighbors if their centers share an edge of the
triangulation~\citep{dea-qua-bir-jua2018}. The pores of an eroded
geometry are illustrated in figure~\ref{fig:Eroding100gap}(a). We do not
consider pores between eroding bodies and the solid wall $\Gamma$, so
some of the grains near the porous region boundary only have two
neighbors. Having defined the pore sizes, we plot its distribution in
figure~\ref{fig:Eroding100gap}(b) and compare it with the Weibull
distribution, a distribution used by others to characterize pore
sizes~\citep{ioa-cha1993}.  In section~\ref{sec:Eroding100}, we
investigate the effect of erosion on the pore size distribution.
\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The pore sizes between
neighboring grains. Two grains are neighbors if they share an edge of
the Delanuay triangulation with nodes at the center of each eroding
grain. (b) The pore sizes of an eroded geometry. The black curve is the
Weibull distribution with the same first two moments as the data.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Consider a single eroding body close to a solid wall with the shear flow
$\UU(\xx) = (y+1,0)$
%\begin{align}
%  \UU(\xx)=\left[
%  \begin{array}{c}
%    y+1 \\ 0
%  \end{array}
%  \right],
%\end{align}
imposed on $\Gamma$. \citet{mit-spa2017} performed a similar
three-dimensional experiment using a second-order quadrature method.
Their initial body is a sphere with its center located $1.5$ radii above
the solid wall.  We initialize the two-dimensional eroding body with
radius $r = 0.4$, and we conduct numerical experiments where the initial
distance between the grain and the solid wall is $h$, $h/2$, and $h/10$,
where $h = 2\upi r/N_\iin$.  If we used the trapezoid rule and required
an error that is comparable to the Barycentric quadrature rule, the body
with an initial distance of $h/10$ from the solid wall would require
$6,400$ discretization points, and the outer wall would require $50,000$
discretization points.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding grains with the Hagen-Poiseuille flow $\UU(\xx) =
U \left(1-y^2,0 \right)$
%\begin{align}
%  \UU(\xx)=U \left[
%  \begin{array}{c}
%    1-y^2 \\ 0
%  \end{array}
%  \right],
%\end{align}
imposed on $\Gamma$. The flow rate $U$ is chosen so that the average
pressure drop from $x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$
once all the grains have vanished.  The vorticity and grain
configuration at four equispaced times are shown in
figure~\ref{fig:Eroding20vort}.  Initially, several of the grains are
closer to the outer wall than the $5h$ threshold required for the
trapezoid rule to achieve machine precision.  In particular, the
distance between bodies 1, 6, 13, and 15 and the outer wall is $1.3h$,
$2.9h$, $2.8h$, and $1.3h$, respectively, where $h$ is the arclength
spacing of the outer wall $\Gamma$.  In addition, the distance between
several pairs of eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and
14 \& 18, is too small to be resolve with the trapezoid rule.  By using
the Barycentric quadrature rule, the interaction between these
nearly-touching bodies is resolved to the desired accuracy, and erosion
can be simulated until all the bodies have vanished.

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} 20 bodies eroding in a
Hagen-Poiseuille flow.  The four snapshots are equispaced in time, and
the color is the fluid vorticity. In the fourth frame, bodies 4, 12, and
17 have vanished, and body 19 has almost vanished.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20densen}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The time-dependent area
fraction of a geometry with 20 eroding bodies. (b) The time-dependent
flow rate, $U$, for a fixed pressure drop across the channel.  The flow
rate is initially small, but it eventually increases as an exponential
law (dashed line) towards the maximum flow rate $U=1$.}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from figure~\ref{fig:Eroding20vort}. The streamlines are
initially equispaced at $(-1,y)$, where $y \in (-1,1)$. The first five
snapshots are equispaced in time.  The bottom right frame is a
magnification of the fifth snapshot, but with additional streamlines.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The local tortuosity of a porous
geometry initialized with 20 grains after eroding to a porosity of
62.9\%.  (a) The $x$-component of the velocity at the inlet,
$u_1(-1,y)$, normalized by its maximum velocity of $2.98 \times
10^{-3}$. (b) The local tortuosity $\tau(y)$ on the cross section $x =
-1$. (c) The streamlines resulting in the ten largest differences of
local tortuosity between neighboring streamlines.  Neighboring
streamlines have the same color.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} 20 bodies eroding in a
Hagen-Poiseuille flow. The snapshots are equispaced in time, and the
color is the fluid vorticity. In addition to the channels that develop
between the bodies and the solid walls, erosion leads to two main
channels through the geometry---one in the top half and one near the
middle.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/100b_50}
  \includegraphics[height=0.227\textwidth, trim=0cm 0cm 2.8cm 0cm, clip=true]
    {./figs/100b_100}
  \includegraphics[height=0.227\textwidth]{./figs/100b_150}
  \includegraphics[height=0.227\textwidth]{./figs/100b_200}
\end{center}
\caption{\label{fig:Eroding100vort} 100 bodies eroding in a
Hagen-Poiseuille flow. The four snapshots are equispaced in time, and
the color is the fluid vorticity.  Erosion creates many channels
connecting the inlet to the outlet.}
\end{figure}

%\begin{figure}
%\begin{subfigure}[b]{0.5\textwidth}
%\includegraphics*[height =0.8\linewidth]{./figs/100b_50}
%\caption{100 bodies, $\phi = 55.54\%$}
%\end{subfigure}%
%\begin{subfigure}[b]{0.5\textwidth}
%\includegraphics*[height =0.8\linewidth]{./figs/100b_100}
%\caption{99 bodies, $\phi = 62.98\%$}
%\end{subfigure}
%\begin{subfigure}[b]{0.5\textwidth}
%\includegraphics*[height =0.8\linewidth]{./figs/100b_150}
%\caption{94 bodies, $\phi = 72.22\%$}
%\end{subfigure}%
%\begin{subfigure}[b]{0.5\textwidth}
%\includegraphics*[height =0.8\linewidth]{./figs/100b_200}
%\caption{82 bodies, $\phi = 83.67\%$}
%\end{subfigure}
%\caption{\label{fig:Eroding100vort} 100 bodies eroding in a
%Hagen-Poiseuille flow. The four snapshots are equispaced in time, and
%the color is the fluid vorticity.  Erosion creates many channels
%connecting the inlet to the outlet.}
%\end{figure}

In figure~\ref{fig:Eroding100Transport}(a), we plot the tortuosity as a
function of the porosity.  The initial geometry has a porosity of $\phi
= 50.09\%$ and the tortuosity is $T = 1.20$.  The tortuosity is computed
with both the length of the streamlines~\eqref{eqn:tortuosity1} (red
stars) and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding100tort}(c).
Again, the two tortuosity formulas give similar results.  For this
geometry, the tortuosity decreases monotonically at almost all
porosities.  However, the tortuosity undergoes a sudden increase near
the end of the simulation, and we have observed this behavior in other
examples.  The increase is caused by a single small body near the middle
of the channel being completely eroded.  While this results in
straighter streamlines, therefore reducing $\lambda$, the horizontal
flow, $u_1(y)$, increases since there is no longer a no-slip boundary,
and this increases the tortuosity.  We also compute the lines of best
fit using the porosity-tortuosity models~\eqref{eqn:tortuosityModels}.
The black dashed line in figure~\ref{fig:Eroding100Transport}(a) is the
line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2459},
\end{align}
with a root-mean-square error of $5.50 \times 10^{-3}$.  We note a
slightly better root-mean-square error of $5.20 \times 10^{-3}$ is
possible with the model
\begin{align}
  \widehat{T}(\phi) = 1 - 0.2631 \ln(\phi).
\end{align}

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}

\caption{\label{fig:Eroding20tort} The local tortuosity of a porous
geometry initialized with 20 grains after eroding to a porosity of
62.9\%.  (a) The $x$-component of the velocity at the inlet,
$u_1(-1,y)$, normalized by its maximum velocity of $2.98 \times
10^{-3}$. (b) The local tortuosity $\tau(y)$ on the cross section $x =
-1$. (c) The streamlines resulting in the ten largest differences of
local tortuosity between neighboring streamlines.  Neighboring
streamlines have the same color.}

\caption{\label{fig:Eroding100tort} The local tortuosity of a porous
geometry initialized with 100 grains after eroding to a porosity of
62.98\%.  (a) The $x$-component of the velocity at the inlet, $u_1(-1,
y)$, normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.
Note that this maximum velocity is about a order of magnitude smaller
than the 20 body example in figure~\ref{fig:Eroding20tort}.  (b) The
local tortuosity $\tau(y)$ on the cross section $x = -1$. Compared to
figure~\ref{fig:Eroding20tort}, this example has more small bodies, and
this results in more discontinuities in the local tortuosity.  (c) The
trajectories of 200 tracers initialized at $x = -1$.}
\end{figure}
